Prospective testing for the prevalence or transience of a shock effect over a set time horizon before it occurs

### 2022-08-06 prevalence_testing/postshocktest2022draft.tex
addition of revised results

In this data example we set $T_i- T_i^* = 30$ and $K = 30$. Note that $K = 30$ is a sensible  training sample size, see discussions for example in  \cite{lin2021minimizing}. The quantity $T_i-T_i^*-H$ corresponds to the post-shock time length that analysts are interested in predicting whether the impact of the shock persists.  We set $T_i- T_i^* = 30$  in order to make the post-shock time length moderately large and also for illustration purpose. For sensitivity analysis, we implemented the analysis for different  $H= 2, 4, 6, 8$.

@@

\begin{table}[H]
	\caption{Results for Conoco Philips ($n=3$), $\mbf{W}^* = (0.221, 0, 0.779)$ from scaled covariates} \label{coptable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$  & Model &   $p_1$ & $\hat{p}_1$ & $|\hat{p}_1-p_1|$ & Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     \multirow{2}{*}{2} & $\mc{M}_0$  &   1 & 0.225  & 0.775 & 0 (Yes) & 1 (No)  \\      
      & $\mc{M}$  & 1 & 1 & $0$   & 0 (Yes) &0 (Yes)   \\[.2cm]
     \multirow{2}{*}{4} & $\mc{M}_0$ &   0 & 0.953  & $0.953$ & 0 (No) &0 (No)    \\
      &  $\mc{M}$ & 1& 1 & 0 &0 (Yes)  &0 (Yes)  \\[.2cm]
     \multirow{2}{*}{6} &  $\mc{M}_0$ &   0   & $1$ & 1 & 0 (No) &0 (No)   \\
      &  $\mc{M}$ & 1 & 1  & 0 &0 (Yes) & 0 (Yes) \\[.2cm]
     \multirow{2}{*}{8} & $\mc{M}_0$ & 0  & 1 & 1 & 0 (No) & 0 (No)\\
     & $\mc{M}$ & 1 & 1 & 0 & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

@@

Let $a_i^1$ and $a_i^2$ denote the AICs of $\mc{M}_0$ and $\mc{M}$ for $i$th time series, respectively. Based on \eqref{model-gen}, which describes $\mc{M}$ and the description of $\mc{M}$ in the beginning of Section \ref{sec::ne}, it may be inferred that $\mc{M}$ captures the post-shock dynamic change whereas $\mc{M}_0$ suggests a persistent post-shock additive location shift. Using the synthetic pairwise model selection procedures specified in Corollary \ref{coro2}, it turns out that $\sum_{i=2}^4 w_i^* I(a_i^1 > a_i^2)=1 >0.5$. Using this aggregated model selection approach, we prefer $\mc{M}$ over $\mc{M}_0$. \textcolor{red}{For reference, the AICs for $\mc{M}_0$ in the donors are $(297.410, 413.357, 295.174)$ whereas those for $\mc{M}$ in the donors are $(302.835, 409.544, 280.609)$. }

From Table \ref{coptable}, we can see that $p_1$ for $\mc{M}$ is not sensitive to the changes of $H$ and is uniformly 1. Under any reasonable significance level, based on the features of $\mc{M}$, we infer that post-shock dynamics do not differ from the pre-shock one and the shock is not persistent. In contrast, $p_1$ of $\mc{M}_0$ equals 1 for $H = 2$ whereas equals 0 for $H = 4, 6, 8$. It is possible that $\mc{M}_0$ does not fit the data well. In practice, a simple additive post-shock location-shift is rare and more complicated structures cannot be explained using this.

Two dashed vertical lines are plotted in Figure \ref{copshocktransience}, where the left one corresponds to $t = T_1^*$ and the right one  corresponds to $t = T_1^*+1$. As shown by \cite{lin2021minimizing}, $y_{1,t}$ experienced a substantial shock at  $t = T_1^*+1$ and it is evident from $y_{1, T_1^*}-y_{1, T_1^*+1}$ shown in Figure 1. We also graphed a benchmark horizontal dashed line for comparison, which corresponds to the stock price $y_{1, T_1^*}$. However, notice that even if $y_{1,t}$ first experienced a considerable drop at $t = T_1^*+1$,   it started to revert back and became closer to the benchmark horizontal line as $t$ increases for $t > T_1^*$. Note that this observation confirms our model selection and what $\mc{M}$ suggests for shock transience. It is also important to re-emphasize that  the test compares the forecasts from $t = T_1^* + H+1$ to $t=T_1$. In other words, the beginning dramatic drop as seen from Figure \ref{copshocktransience} around March is not considered in the forecast comparison. It is reasonable to our objective as we are interested in long-term persistence rather-than short-term substantial changes.

@@

\textcolor{red}{According to the last two columns of Table \ref{coptable}, the decisions yielded by voting and weighted voting are correct for $\mc{M}$ and $H = 2, 4, 6, 8$. However, the decisions for voting and weighted voting are incorrect for $\mc{M}_0$ and $H=2, 4, 6, 8$ except for voting for $\mc{M}_0$ at $H = 2$. The interpretations for this are similar to those when we analyze $|\hat{p}_1-p_1|$. Note that the decisions using $\hat{p}_1$ and weighted voting are similar. It is reasonable as they are functions of $\mbf{W}^*$. Because $p_1$ is uniformly 1 for $\mc{M}$, it is reasonable to expect $\P(p_1\leq .05)$ be small. Note that Proposition \ref{votprop} and Corollary \ref{coro2} show that the decisions made by voting and weighted voting procedures should be correct with very high probability if we assume the asymptotic approximation of the expected absolute error works well for  $n = 3$.}

@@

\begin{table}[H]
	\caption{Results for Month-on-Month Change in Log Nonfarm Payrolls (Seasonally Adjusted) ($n=3$), $\mbf{W}^* = (0.196, 0.804, 0.000)$ from scaled covariates}   \label{uemtable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$  & Model &   $p_1$ & $\hat{p}_1$ & $|\hat{p}_1-p_1|$ & Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     \multirow{2}{*}{2} & $\mc{M}_0$  &   0.450 & 0.511  & 0.061 & 0 (Yes) & 0 (Yes)\\      
      & $\mc{M}$  & 0.240 & 0.583 & $0.343$   & 0 (Yes) &0 (Yes)   \\[.2cm]
     \multirow{2}{*}{3} & $\mc{M}_0$ &   0.500 & 0.469  & $0.031$ & 0 (Yes) & 0 (Yes)    \\
      &  $\mc{M}$ & 1 & 0.576 & 0.424 & 0 (Yes) & 0 (Yes)  \\[.2cm]
     \multirow{2}{*}{4} &  $\mc{M}_0$ &   0.660   & $0.443$ & 0.217 & 0 (Yes) & 0 (Yes)  \\
      &  $\mc{M}$ & 1.000 & 0.494  & 0.506 & 0 (Yes) & 0 (Yes) \\[.2cm]
     \multirow{2}{*}{5} & $\mc{M}_0$ & 0.700  & 0.556 & 0.144 & 0 (Yes) & 0 (Yes)\\
     & $\mc{M}$ & 0.470 & 1.000 & 0.530 & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

@@

\begin{align*}
	\phi_{i,j}, \theta_{i,j}  & \simiid U(0,1) \\ 
	\delta_{i,j} & \simiid N(1,0.5^2) \\
	\alpha_i & \simiid N(10, 2) \\
	T_i^* & \simiid \min\Big(90, \big[Gamma(15,10)\big]\Big) \\
	\varepsilon_{i,t} & \simiid N(0,1)\\
	X_{i,j} & \simiid Gamma(1,2) \\
	\tilde{\theta}_{i,j},\tilde{\phi}_{i,j} & \simiid Beta(\|\mbf{x}_{i,T_i^*+1}\|,1)
\end{align*}
Vary $L = 14, 30, 45,60$.

>>>>

In this data example we set $T_i- T_i^* = 50$ and $K = 30$. Note that $K = 30$ is a sensible  training sample size, see discussions for example in  \cite{lin2021minimizing}. The quantity $T_i-T_i^*-\ceil{1.5 \cdot \sqrt{T_i-T_i^*}}$ corresponds to the post-shock time length that analysts are interested in predicting whether the impact of the shock persists.  We set $T_i- T_i^* = 50$  in order to make the post-shock time length moderately large and also for illustration purpose. For sensitivity analysis, we implemented the analysis for different  $H= 2, 4, 6, 8$.

@@

\begin{table}[H]
	\caption{Results for Conoco Philips ($n=3$), $\mbf{W}^* = (0.221, 0, 0.779)$ from scaled covariates} \label{coptable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$  & Model &   $p_1$ & $\hat{p}_1$ & $|\hat{p}_1-p_1|$ & Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     \multirow{2}{*}{2} & $\mc{M}_0$  &   1 & 0.023  & 0.977 & 0 (Yes) & 1 (No)  \\      
      & $\mc{M}$  & 1 & 1 & $0$   & 0 (Yes) &0 (Yes)   \\[.2cm]
     \multirow{2}{*}{4} & $\mc{M}_0$ &   1 & 0.125 & $0.875$ & 0 (Yes) &1 (No)    \\
      &  $\mc{M}$ & 1& 1 & 0 &0 (Yes)  &0 (Yes)  \\[.2cm]
     \multirow{2}{*}{6} &  $\mc{M}_0$ &   0   & $0.999$ & 0.999 & 0 (No) &0 (No)   \\
      &  $\mc{M}$ & 1 & 1  & 0 &0 (Yes) & 0 (Yes) \\[.2cm]
     \multirow{2}{*}{8} & $\mc{M}_0$ & 0  & 1 & 1 & 0 (No) & 0 (No)\\
     & $\mc{M}$ & 1 & 1 & 0 & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

@@

Let $a_i^1$ and $a_i^2$ denote the AICs of $\mc{M}_0$ and $\mc{M}$ for $i$th time series, respectively. Based on \eqref{model-gen}, which describes $\mc{M}$ and the description of $\mc{M}$ in the beginning of Section \ref{sec::ne}, it may be inferred that $\mc{M}$ captures the post-shock dynamic change whereas $\mc{M}_0$ suggests a persistent post-shock additive location shift. Using the synthetic pairwise model selection procedures specified in Corollary \ref{coro2}, it turns out that $\sum_{i=2}^4 w_i^* I(a_i^1 > a_i^2)=0.779 >0.5$. Using this aggregated model selection approach, we prefer $\mc{M}$ over $\mc{M}_0$. \textcolor{red}{For reference, the AICs for $\mc{M}_0$ in the donors are $(348.261, 504.887, 377.073)$ whereas those for $\mc{M}$ in the donors are $(349.784, 505.012, 364.108)$. }

From Table \ref{coptable}, we can see that $p_1$ for $\mc{M}$ is not sensitive to the changes of $H$ and is uniformly 1. Under any reasonable significance level, based on the features of $\mc{M}$, we infer that post-shock dynamics do not differ from the pre-shock one and the shock is not persistent. In contrast, $p_1$ of $\mc{M}_0$ equals 1 for $H = 2, 4$ whereas equals 0 for $H = 6, 8$. It is possible that $\mc{M}_0$ does not fit the data well. In practice, a simple additive post-shock location-shift is rare and more complicated structures cannot be explained using this.

Two dashed vertical lines are plotted in Figure \ref{copshocktransience}, where the left one corresponds to $t = T_1^*$ and the right one  corresponds to $t = T_1^*+1$. As shown by \cite{lin2021minimizing}, $y_{1,t}$ experienced a substantial shock at  $t = T_1^*+1$ and it is evident from $y_{1, T_1^*}-y_{1, T_1^*+1}$ shown in Figure 1. We also graphed a benchmark horizontal dashed line for comparison, which corresponds to the stock price $y_{1, T_1^*}$. However, notice that even if $y_{1,t}$ first experienced a considerable drop at $t = T_1^*+1$,   it started to revert back and became closer to the benchmark horizontal line as $t$ increases for $t > T_1^*$. Note that this observation confirms our model selection and what $\mc{M}$ suggests for shock transience. It is also important to re-emphasize that  the test compares the forecasts from $t = T_1^* + H+\ceil{1.5 \cdot \sqrt{T_1-T_1^*}}$ to $t=T_1$. In other words, the beginning dramatic drop as seen from Figure \ref{copshocktransience} around March is not considered in the forecast comparison. It is reasonable to our objective as we are interested in long-term persistence rather-than short-term substantial changes.

@@

\textcolor{red}{According to the last two columns of Table \ref{coptable}, the decisions yielded by voting and weighted voting are correct for $\mc{M}$ and $H = 2, 4, 6, 8$. However, the decisions for voting and weighted voting are incorrect for $\mc{M}_0$ and $H=2, 4, 6, 8$ except for voting for $\mc{M}_0$ at $H = 2,4$. Here, it is worth noting that weighted voting and voting may differ due to $\mbf{W}^*$. Such difference may be valuable if the weight is concentrated in one donor that does not tell a consistent story to that of time series of interest. Next, the interpretations for these results are similar to those when we analyze $|\hat{p}_1-p_1|$. Note that the decisions using $\hat{p}_1$ and weighted voting are similar. It is reasonable as they are functions of $\mbf{W}^*$. Because $p_1$ is uniformly 1 for $\mc{M}$, it is reasonable to expect $\P(p_1\leq .05)$ be small. Note that Proposition \ref{votprop} and Corollary \ref{coro2} show that the decisions made by voting and weighted voting procedures should be correct with very high probability if we assume the asymptotic approximation of the expected absolute error works well for  $n = 3$.}

@@

\begin{table}[H]
	\caption{Results for Month-on-Month Change in Log Nonfarm Payrolls (Seasonally Adjusted) ($n=3$), $\mbf{W}^* = (0.196, 0.804, 0.000)$ from scaled covariates}   \label{uemtable}
\begin{center}
  \begin{tabular}{cccccccccr}
     $H$  & Model &   $p_1$ & $\hat{p}_1$ & $|\hat{p}_1-p_1|$ & Vote (Correct?) & weighted Vote (Correct?)   \\
     \hline 
     \multirow{2}{*}{2} & $\mc{M}_0$  &   0.300 & 0.443  & 0.143 & 0 (Yes) & 0 (Yes)\\      
      & $\mc{M}$  & 0.775 & 0.591 & $0.184$   & 0 (Yes) &0 (Yes)   \\[.2cm]
     \multirow{2}{*}{3} & $\mc{M}_0$ &   0.215 & 0.477 & $0.262$ & 0 (Yes) & 0 (Yes)    \\
      &  $\mc{M}$ & 0.740 & 0.600 & 0.140 & 0 (Yes) & 0 (Yes)  \\[.2cm]
     \multirow{2}{*}{4} &  $\mc{M}_0$ &   0.615   & $0.343$ & 0.272 & 0 (Yes) & 0 (Yes)  \\
      &  $\mc{M}$ & 0.780 & 0.464  & 0.316 & 0 (Yes) & 0 (Yes) \\[.2cm]
     \multirow{2}{*}{5} & $\mc{M}_0$ & 0.030 & 0.377 & 0.347 & 0 (No) & 0 (No)\\
     & $\mc{M}$ & 0.940 & 0.293 & 0.647 & 0 (Yes) & 0 (Yes)
  \end{tabular}
\end{center}	
\end{table}

@@

\begin{align*}
	\phi_{i,j}, \theta_{i,j}  & \simiid U(-0.4,0.4) \\ 
	\delta_{i,j} & \simiid N(1,0.5^2) \\
	\alpha_i & \simiid N(10, 2) \\
	T_i^* & \simiid \min\Big(90, \big[Gamma(15,10)\big]\Big) \\
	\varepsilon_{i,t} & \simiid N(0,1)\\
	X_{i,j} & \simiid Gamma(1,c) \\
	e_{i,j} & \simiid beta(\alpha, \beta) \\
	\tilde{\theta}_{i,j}&= \theta_{i,j} + \frac{1}{5}\cdot \Big(e_{ij}-\frac{1}{2}\Big)
\end{align*}
